\begin{thebibliography}{99.}

\bibitem{adams_2003_2}\aut{Adams, R.; Fournier J.} \nob{Sobolev Spaces.} 2 ed. Kidlington: Academic Press, 2003.

\bibitem{agranovitch_2015_1}\aut{Agranovich, M. S.} \nob{Sobolev Spaces, Their Generalizations, and Elliptic Problems
in Smooth and Lipschitz Domains.} 1 ed. London: Springer-Verlag, 2015.

\bibitem{appleby_1987}\aut{Appleby, P. G. et al} \textit{On the Classification of Isotropic Tensors} in \nob{Glasgow Mathematical Journal}, Vol. 29, No. 2. Cambridge: Cambridge University Press, 1987, pp. 185-196.

\bibitem{artzy_1965_1}\aut{Artzy, R.} \nob{Linear Geometry.} 1 ed. Reading: Addison-Wesley Publishing Company, 1965.

\bibitem{backus_1997_1}\aut{Backus, G.} \nob{Continuum Mechanics.} Available at
<\url{http:// samizdat.mines.edu/backus/}>, Accessed on feb 2004.

\bibitem{bishop_1980_1}\aut{Bishop, R. L.; Goldberg, S. I.} \nob{Tensor Analysis on Manifolds.} 1 ed. Mineola: Dover Publications, 1980.

\bibitem{bowen_2008_1}\aut{Bowen, R. M.; Wang, C. -C.} \nob{Introduction to Vectors and Tensors.} 1 ed. Mineola: Dover Publications, 2008.

\bibitem{cameron_1999_1}\aut{Cameron, P. J.} \nob{Sets, Logic and Categories.} 1 ed. London: Springer-Verlag, 1999.

\bibitem{cervantes_2010}\aut{Cervantes, M.} \nob{Don Quijote De La Mancha} 1 ed. Madrid: Castalia, 2010.

\bibitem{chadwick_1999_2}\aut{Chadwick, P.} \nob{Continuum Mechanics: Concise Theory and Problems.} 2 ed. Mineola: Dover Publications, 1999.

\bibitem{christoffel_1869_1}\aut{Christoffel, E. B.} \textit{Ueber die Transformation der homogenen Differentialausdrücke zweiten Grades} in \nob{Journal für die Reine und Angewandte Mathematik.} Vol. 70, 1869, pp. 46-70.

\bibitem{ciarlet_1988_2}\aut{Ciarlet, P. G.} \nob{Mathematical Elasticity - Volume 1:
Three Dimensional Elasticity.} 1. ed. Amsterdam: Elsevier Science Publishers BV, 1988.

\bibitem{connel_1999_1}\aut{Connel, E. H.} \nob{Elements of Abstract and Linear Algebra.}
Available at <\url{http://www.math.miami.edu/~ec/book/}>, Accessed on feb 2004.

\bibitem{conrad_XXXX_1}\aut{Conrad, F.} \nob{Tensor Products.} Available at
<\url{http://www.math.uconn.edu/~kconrad/blurbs/linmultialg/tensorprod.pdf}>, Accessed on sept 2017.

\bibitem{derbyshire_2006_1}\aut{Derbyshire, J.} \nob{Unknown Quantity: a Real and Imaginary History of Algebra.}
1 ed. Washington,D.C.: Joseph Henry Press, 2006.

\bibitem{dieudonne_1969_1}\aut{Dieudonné, J.} \nob{Foundations Of Modern Analysis.}
1 ed. New York: Academic Press, 1969.

\bibitem{docarmo_2016}\aut{Do Carmo, M. P.} \nob{Differential Geometry of Curves and Surfaces.}
2 ed. Mineola: Dover Publications, 2016.

\bibitem{dodson_1991}\aut{Dodson, C.T.J.; Poston, T.} \nob{Tensor Geomery: The Geometric Viewpoint and its Uses.}
1 ed. Berlin: Springer-Verlag Inc., 1991.

\bibitem{einstein_1915_1}\aut{Einstein, A.} \textit{Zur allgemeinen Relativitätstheorie} in \nob{Preussische Akademie der Wissenschaften, Sitzungsberichte.} Berlim, 1915, pp. 778-786, 799-801.

\bibitem{figueiredo_1964_1}\aut{Figueiredo, D. G.} \textit{A Simplified Proof of the Divergence Theorem} in \nob{The American Mathematical Monthly}, Vol. 71, No. 6, 1964, pp. 619-622.

\bibitem{flanders_1989_1}\aut{Flanders, H.} \nob{Differential Forms With the Applications to the Physical Sciences.} 1 ed. Mineola: Dover Publications, 1989.

\bibitem{gibbs_1881}\aut{Gibbs, J. W.} \nob{Elements of Vector Analysis Arranged for the Use of Students in Physics.} 1. ed.  New Haven: Tuttle, Morehouse \& Taylor, 1881.

\bibitem{gurtin_1981}\aut{Gurtin, M. E.} \nob{An Introduction to Continuum Mechanics.} 1. ed. San Diego: Academic Press, 1981.

\bibitem{gurtin_2010}\aut{Gurtin, M. E. et al.} \nob{The Mechanics and Thermodynamics of Continua.} 1. ed. Cambridge: Cambridge University Press, 2010.

\bibitem{halmos_1974_1}\aut{Halmos, P. R.} \nob{Measure Theory.} 1 ed. New York: Springer-Verlag Inc., 1974.

\bibitem{halmos_19742_1}\aut{Halmos, P. R.} \nob{Naive Set Theory.} 1 ed. New York: Springer-Verlag Inc., 1974.

\bibitem{hamilton_1846_1}\aut{Hamilton, W. R.} \textit{On Quaternions; or on a new System of Imaginaries in Algebra} in \nob{ The London, Edinburgh and Dublin Philosophical Magazine and Journal of Science: Volume XXIX.} 1846, pp. 26-31.

\bibitem{hermann_1975}\aut{Hermann, R.} \nob{Ricci and Levi-Civita's Tensor Analysis Paper.} 1 ed. Brookline: Math Sci Press, 1975.

\bibitem{hermes_1974_1}\aut{Hermes, H.; Markwald, W.} \textit{Foundations of Mathematics} in \nob{ Fundamentals of Mathematics: Volume I.} 1 ed. Massachusetts: MIT Press, 1987, pp. 1-86.

\bibitem{johnson_1970}\aut{Johnson, C. R.} \textit{Positive Definite Matrices} in \nob{The American Mathematical Monthly} Vol. 77, No. 3. New York: Mathematical Assossiation of America, 1970, pp. 259-264.

\bibitem{knuth_1997_1}\aut{Knuth, D. E.} \nob{The Art of Computer Programming, Volume 1.} 1 ed. Boston: Addison-Wesley, 1997.

\bibitem{kreyszig_1978_1}\aut{Kreyszig, E.} \nob{Introductory Functional Analysis With Applications.} 1 ed. Boston: John Wiley \& Sons, 1978.

\bibitem{lanczos_1970_1}\aut{Lanczos, C.} \nob{Space Through The Ages: The Evolution of Geometrical Ideas From Pythagoras to Hilbert and Einstein.} 1 ed. London: Academic Press, 1970.

\bibitem{lang_1993_3}\aut{Lang, S.} \nob{Real and Functional Analysis.} 3 ed. Berlin: Springer-Verlag, 1993.

\bibitem{levi_2013_1}\aut{Levi-Civita, T.} \nob{The Absolute Differential Calculus (Calculus of Tensors).} 1 ed. Mineola: Dover Publications, 2013.

\bibitem{loomis_2014_1}\aut{Loomis, L. H.; Sternberg, S.} \nob{Advanced Calculus.} 1
ed. Hackensack: World Scientific Publishing Co., 2014.

\bibitem{luque_2006_1}\aut{Luque, J.; Thibon, J.} \nob{Hyperdeterminantal Calculations of Selberg's And Aomoto's Integrals.} Available at <\url{https://arxiv.org/pdf/math/0607410.pdf}>, Accessed on july 2018.

\bibitem{malament_2005_1}\aut{Malament, D. B.} \nob{Notes on Geometry and Spacetime.}
Available at
<\url{http://www.lps.uci.edu/home/fac-staff/faculty/malament/geometryspacetimedocs/}>,
Accessed on aug 2006.

\bibitem{marcus_1973_1}\aut{Marcus, M.} \nob{Finite Dimensional Multilinear Algebra, Part I.} 1 ed. New York: Marcel Dekker Inc., 1973.

\bibitem{milne_2003_1}\aut{Milne J. S.} \nob{Group Theory.} Available at
<\url{http:// www.jmilne.org/math/CourseNotes/}>, Accessed on feb 2004.

\bibitem{munroe_1971_1}\aut{Munroe, M. E.} \nob{Measure and Integration} 1 ed.
Reading: Addison-Wesley Publishing Company, 1971.

\bibitem{ogden_1997}\aut{Ogden, R.W.} \nob{Non-Linear Elastic Deformations.} 1 ed. Mineola: Dover Publications, 1997.

\bibitem{ricci_1900_1}\aut{Ricci, G.; Levi-Civita, T.} \textit{Méthodes de Calcul Différentiel Absolu et Leurs Applications} in \nob{Mathematische Annalen.} Vol 54, No. 1-2. Berlin: Springer, 1900, pp. 125-201.

\bibitem{segel_1999_1}\aut{Segel, L.} \nob{Mathematics Applied to Continuum Mechanics.} 1 ed. New York: Macmillan Publishing Co., 1977.

\bibitem{shen_2002_1}\aut{Shen, A.; Vereshchagin, N. K.} \nob{Basic Set Theory} 1 ed. Providence:  American Mathematical Society, 2002.

\bibitem{snapper_1971_1}\aut{Snapper, E.; Troyer, R. J.} \nob{Metric Affine Geometry.} 1
ed. New York: Academic Press, 1971.

\bibitem{strang_2006_4}\aut{Strang, G.} \nob{Linear Algebra and Its Applications.} 4 ed. Pacific Grove: Brooks-Cole Publishing, 2006.

\bibitem{soutasLittle_1999_1}\aut{Soutas-Little, R. W.} \nob{Elasticity.} 1 ed. Mineola: Dover Publications, 1999.

\bibitem{spivak_1965_1}\aut{Spivak, M.} \nob{Calculus on Manifolds: a Modern Approach to Classical Theorems of Advanced Calculus} 1 ed. Princeton: Perseus Books Publishing, 1965.

\bibitem{truesdell_1992}\aut{Truesdell, C. A.; Noll, W.} \nob{The Non-Linear Field Theories of Mechanics.} 1 ed. Berlin: Springer-Verlag Inc., 1992.

\bibitem{vaisman_1980_1}\aut{Vaisman, I.} \nob{Foundations of Three Dimensional Euclidean
Geometry.} New York: Marcel Dekker Inc., 1980.

\bibitem{vinberg_2003_1}\aut{Vinberg, E. B.} \nob{A Course in Algebra.} Rhode Island: American Mathematical Society, 2003.

\bibitem{voigt_1898_1}\aut{Voigt, W.} \nob{Die Fundamentalen Physikalischen Eigenschaften der Krystalle in Elementarer Darstellung.} Leipzig: Verlag Von Veit \& Comp., 1898.

\bibitem{waerden_1985_1}\aut{Waerden, B.} \nob{A History of Algebra: From al-Khw\=arizm\={\i} to Emmy Noether} 1 ed. Berlin: Springer-Verlag Inc., 1985.

\bibitem{waleffe_1997_1}\aut{Waleffe, F.} \nob{Tensor Product and Tensors.} Available at
<\url{http://www.math.wisc.edu/~milewski/}>, Accessed on october 2004.

\bibitem{weil_1953_2}\aut{Weyl, H.} \nob{The Classical Groups: Their Invariants and Representations.} 2 ed. Princeton: Princeton University Press, 1953.

\bibitem{wouk_1979_1}\aut{Wouk, A.} \nob{A Course of Applied Functional Analysis.} 1 ed.
New York: John Wiley \& Sons Inc., 1979.

\bibitem{zeidler_1995_1}\aut{Zeidler, E.} \nob{Applied Functional Analysis: Main Principles and Their Applications.}
1 ed. New York: Springer-Verlag Inc., 1995.

\end{thebibliography}
