\begin{thebibliography}{99.}

\bibitem{backus_1997_2}\aut{Backus, G.} \nob{ Continuum Mechanics.} Available at
<\url{http:// samizdat.mines.edu/backus/}>, Accessed on feb 2004.

\bibitem{benvenuto_1991}\aut{Benvenuto, E.} \nob{An Introduction to the History of Structural Mechanics, Part I: Statics and Resistance of Solids}. 1. ed. New York: Springer-Verlag, 1991.

\bibitem{jakob_1694}\aut{Bernoulli, Jakob.} \textit{Curvatura Laminae Elasticae. Ejus Identitas cum Curvatura Lintei a pondere inclusi fluidi expansi, \&c.} in \nob{Acta Eruditorum}, 1694, pp. 262-276.

\bibitem{jakob_1695}\aut{Bernoulli, Jakob.} \textit{Explicationes, annotationes et additiones ad ea, quae in actis sup. anni de curva elastica, isochrona paracentrica, et velaria, \&c.} in \nob{Acta Eruditorum}, 1695, pp. 537-553.

\bibitem{belhoste_1991_1}\aut{Belhoste, B.} \nob{Augustin-Louis Cauchy: A Biography.} 1. ed. Berlim: Springer-Verlag, 1991.

\bibitem{bradley_2009}\aut{Bradley, R.; Sandifer, C.} \nob{Cauchy's Cours d'analyse: An Annotated Translation.} 1. ed. New York: Springer-Verlag, 2009.

\bibitem{brandon_2003_2}\aut{Brandon, R.} \nob{ Large Deformation Kinematics.} Available at
<\url{http://www.me.unm.edu/~rmbrann/gobag.html}>, Accessed on jan 2007.

\bibitem{calinger_2016_1}\aut{Calinger, R.} \nob{Leonhard Euler: Mathematical Genius in the Enlightenment.}
1. ed. Princeton: Princeton University Press, 2016.

\bibitem{cauchy_1823}\aut{Cauchy, A.} \textit{Recherches sur L'équilibre et Lemouvement Int\'erieur des Corps Solides ou Fluides, \'Elastiques ou Non \'Elastiques} in \nob{Bulletin Des Sciences de La Société Philomatique
de Paris}, January, 1823, pp. 9-13.

\bibitem{cauchy_1827}\aut{Cauchy, A.} \textit{Sur La Pression ou Tension Dans Les Corps Solides} in \nob{Exercices de Math\'ematiques}, March, 1827, pp. 42-57.

\bibitem{cauchy_1889}\aut{Cauchy, A.} \textit{Sur Quelques Propri\'et\'es des Poly\`edres} in \nob{\OE uvres Compl\'etes d'Augustin Cauchy}, S\'erie II, Tome VII, 1889, pp. 55-59.

\bibitem{cauchySolid_1889}\aut{Cauchy, A.} \textit{De La Pression ou Tension Dans Un Corps Solide} in \nob{\OE uvres Compl\'etes d'Augustin Cauchy}, S\'erie II, Tome VII, 1889, pp. 60-81.

\bibitem{chadwick_1999_1}\aut{Chadwick, P.} \nob{ Continuum Mechanics: Concise Theory and Problems.}
1. ed. Mineola: Dover Publications, 1999.

\bibitem{ciarlet_1988_2_2}\aut{Ciarlet, P. G.} \nob{ Mathematical Elasticity - Volume 1:
Three Dimensional Elasticity.} 1. ed. Amsterdam: Elsevier Science Publishers BV, 1988.


\bibitem{corben_1994_2}\aut{Corben, H. C.; Stehle, P.} \nob{ Classical Mechanics.} 2. ed. New York: Dover Publications, 1994.

\bibitem{dugas_1988_1}\aut{Dugas, R.} \nob{A History of Mechanics.} 1. ed. New York: Dover Publications, 1988.

\bibitem{euler_1749_1}\aut{Euler, L.} \nob{ Scientia Navalis, Volume 1.} 1. ed. St. Petersburg: Academy of Sciences Press, 1749.

\bibitem{fellmann_2007_1}\aut{Fellmann, E.} \nob{Leonhard Euler.} 1. ed. Basel: Birkh\"auser Verlag, 2007.

\bibitem{galileo_1954_2}\aut{Galileo Galilei.} \nob{Dialogs Concerning Two New Sciences.} 1. ed. New York: Dover Publications, 1954.


\bibitem{gurtin_1981_2}\aut{Gurtin, M. E.} \nob{ An Introduction to Continuum Mechanics.} 1. ed. San Diego: Academic Press, 1981.

\bibitem{hanyga_1985_2}\aut{Hanyga, A. } \nob{ Mathematical Theory of Non-Linear Elasticity.} Translated by
R. W. Ogden. 1. ed. Chichester/New York/Warsaw: Elis Horwood Limited/John Wiley \&
Sons/PWN-Polish Scientific Publishers, 1985.

\bibitem{haupt_2002_2}\aut{Haupt, P.} \nob{ Continuum Mechanics and Theory of Materials.} 2. ed. Berlim: Springer-Verlag, 2002.

\bibitem{heilbron_2010_1}\aut{Heilbron, J.L.} \nob{Galileo.} 1. ed. New York: Oxford University Press, 2010.

\bibitem{higdon_1981_3}\aut{Higdon, A. et al.} \nob{Mecânica dos Materiais.} 3. ed. Rio de Janeiro: Guanabara Dois, 1981.

\bibitem{hill_1978_2}\aut{Hill, R.} Aspects of Invariance in Solid Mechanics. \nob{ Advances in Applied Mechanics.} New York: Academic Press, vol. 18, pp. 1-75, 1978.

\bibitem{hooke_1678_1}\aut{Hooke, R.} \nob{Lectures \emph{de Potentia Restitutiva}, Or of Spring: Explaining the Power of Springing Bodies.} 1. ed. London: John Martyn, 1678.

\bibitem{hooke_1705_1}\aut{Hooke, R.} \nob{The Posthumous Works of Robert Hooke.} 1. ed. London: Richard Waller, 1705.


\bibitem{love_1944_2}\aut{Love, A. E. H.} \nob{ A Treatise on the Mathematical Theory of Elasticity.} 4. ed. New York: Dover Publications, 1944.

\bibitem{lubarda_2002_2}\aut{Lubarda, V. A.} \nob{ Elastoplasticity Theory.} 1. ed. Boca Raton: CRC Press,   2002.

\bibitem{lund_2000_1}\aut{Lund, J.; Byrne, J.} \textit{Leonardo Da Vinci's Tensile Strength Tests: Implications For The Discovery Of Engineering Mechanics} in \nob{Civil. Eng. and Env. Syst.} Vol. 00, 2000, pp. 1-8.


\bibitem{malvern_1969_2}\aut{Malvern, L. E.} \nob{ Introduction to the Mechanics of a Continuous Medium.}
1. ed. Englewood Cliffs: Prentice-Hall, 1969.

\bibitem{mariotte_1740_1}\aut{Mariotte, E.} \nob{Traité du mouvement des eaux et des autres corps fluides.} 2. ed. Paris: Jean Néaulme, 1700.

\bibitem{marsden_1994_2}\aut{Marsden, J. E.; Hughes, T. J. R.} \nob{ Mathematical
Foundations of Elasticity.} 1. ed. Mineola: Dover Publications, 1994.


\bibitem{maugin_1993_2}\aut{Maugin, G.} \nob{Material Inhomogeneities in Elasticity.} 1. ed. New York: Chapman \& Hall, 1993.

\bibitem{maugin_2013}\aut{Maugin, G.} \nob{Continuum Mechanics Through the Twentieth Century: A Concise Historical Perspective.} 1. ed. Dordrecht: Springer, 2013.

\bibitem{maugin_2014}\aut{Maugin, G.} \nob{Continuum Mechanics Through the Eighteenth and Nineteenth Centuries.} 1. ed. Switzerland: Springer, 2014.

\bibitem{maugin_2016}\aut{Maugin, G.} \nob{Continuum Mechanics Through the Ages - From the Renaissance to the Twentieth Century.} 1. ed. Switzerland: Springer, 2016.

\bibitem{maxwell_1991_2}\aut{Maxwell, J. C.} \nob{ Matter and Motion.} 1. ed. New York: Dover Publications, 1991.

\bibitem{mikhailov_2005}\aut{Mikhailov, G.} \textit{Daniel Bernoulli, Hydrodynamica (1738)} in \nob{Landmark Writings in Western Mathematics 1640-1940.} 1. ed. Amsterdam: Elsevier Science, 2005, pp. 131-142.

\bibitem{newton_1999_1}\aut{Newton, I.} \nob{The Principia: Mathematical Principles of Natural Philosophy.}, translated by I. Bernard Cohen and Anne Whitman, 1. ed. Berkeley: University of California Press, 1999.

\bibitem{noll_1974_2}\aut{Noll, W.} \nob{ The Foundations of Mechanics and Thermodynamics: Selected Papers.} 1. ed. Berlim: Springer-Verlag, 1974.


\bibitem{ogden_1997_2}\aut{Ogden, R. W.} \nob{ Non-Linear Elastic Deformations.} 1. ed. Mineola: Dover Publications, 1997.

\bibitem{oldfather_1933_1}\aut{Oldfather, W. et al.} \textit{Leonhard Euler's Elastic Curves} in \nob{Isis}, Vol. 20, No. 1,  Chicago: The University of Chicago Press, 1933, pp. 72-160.

\bibitem{podio_2000_2}\aut{Podio-Guidugli, P.} A Primer on Elasticity. \nob{ Journal of Elasticity.} Berlim: Springer-Verlag, vol. 40, pp. 1-104, 2000.

\bibitem{popov_1990_1}\aut{Popov, E.} \nob{ Engineering Mechanics of Solids.} 1. ed. Englewood Cliffs: Prentice Hall, 1990.

\bibitem{prager_1997_2}\aut{Prager, W.} \nob{ Introduction to Mechanics of Continua.} 1. ed. Mineola: Dover Publications, 2004.

\bibitem{sedov_1997_2}\aut{Sedov, L. I.} \nob{ Mechanics of Continuous Media - Volume 1.} Translated by J. P. Nowacki. 1. ed. Farrer Road: World Scientific Publishing Co., 1997.

\bibitem{silhavy_1997_2}\aut{\v{S}ilhavý, M.} \nob{ The Mechanics and Thermomechanics of Continuous Media.} 1. ed. Berlim: Springer-Verlag, 1997.

\bibitem{sokolnikoff_1983_2}\aut{Sokolnikoff, I. S.} \nob{ Mathematical Theory of Elasticity.}
2. ed. Malabar: Robert Krieger Publishing Company, 1983.

\bibitem{talpaert_2002_2}\aut{Talpaert, Y. R.} \nob{ Tensor Analysis and Continuum Mechanics.} 1. ed. Dordrecht: Kluwer Academic Publishers, 2002.

\bibitem{teeman_2005_2}\aut{Temam, R. M.; Miranville, A. M.} \nob{ Mathematical Modeling in Continuum Mechanics.} 2. ed. Cambridge: Cambridge University Press, 2005.

\bibitem{timoshenko_1983_1}\aut{Timoshenko, S.} \nob{History of Strength of Materials.} 1 ed. Mineola: Dover Publications, 1983.

\bibitem{tolstoi_2014_1}\aut{Tolstoy, L.} \nob{War \& Peace.} 1 ed. London: The Folio Society Ltd, 2014.

\bibitem{truesdell_1954_1}\aut{Truesdell, C. A.} \textit{Rational fluid mechanics, 1687-1765: Editor's introduction to vol. II 12 of Euler's works} in \nob{Euleri Opera Omnia, Series II, vol. 12.} 1. ed. Zurich: Füssli, 1954, pp. 1-125.

\bibitem{truesdell_1960}\aut{Truesdell, C. A.} \nob{The Rational Mechanics of Flexible or Elastic Bodies: 1638-1788}. 1. ed. Basel: Birkhäuser, 1960.

\bibitem{truesdell_1968}\aut{Truesdell, C. A.} \nob{Essays in the History of Mechanics}. 1. ed. New York: Springer-Verlag, 1968.

\bibitem{truesdell_1977_2}\aut{Truesdell, C. A.} \nob{ A First Course in Rational Continuum Mechanics - Volume 1}. 1. ed. New York: Academic Press, 1977.


\bibitem{truesdellNoll_1997_2}\aut{Truesdell, C. A.; Noll, W.} \nob{ The Non-Linear Field Theories
of Mechanics.} 3. ed. New York: Springer-Verlag, 2003.

\bibitem{wallace_1998_1}\aut{Wallace, W.} \textit{Galileo's Pisan studies in science and philosophy} in \nob{ The Cambridge Companion to Galileo.} 1. ed. Cambridge: Cambridge University Press, 1998, pp. 27-52.


\bibitem{wangTruesdell_1973_2}\aut{Wang, C.-C; Truesdell, C. A.} \nob{ Introduction to Rational Elasticity.} 1. ed. Leyden: Noordhoff International Publishing, 1973.


\bibitem{weyl_1952_2}\aut{Weyl, H.} \nob{ Space Time Matter.} 4. ed. Mineola: Dover Publications, 1952.

\bibitem{zammattio_1980}\aut{Zammattio, C. et al.} \nob{Leonardo The Scientist.} 1. ed. New York: McGraw-Hill, 1980.



\end{thebibliography}
