\chapter{Continuum Kinematics}

bla, bla, bla, bla


\section{Continuum Mechanics: a Mathematical Model}

In classical terms, Mechanics is the physical science of \textsb{motion}\index{motion} concerned with its descriptive and predictive aspects. The descriptive study of motion, or Kinematics, deals mainly with its geometric measurements while the predictive study, Dynamics, considers its causes. For both of these aspects, the notion of \textsb{time}\index{time} is fundamental, enabling the very idea of motion and most of its related concepts, particularly in Kinematics. For the purposes of Dynamics, \textsb{force}\index{force} is the main concept, being the motive of geometric changes in time, a causality relation, called the \textsb{Laws of Motion}\index{motion! Laws of}, which aims to be as generic as possible and independent of material specificities. But when the influence of these specificities can not be disregarded for the mechanical phenomena under study, particular \textsb{constitutive}\index{description!constitutive} relations have to be considered. 


In a restricted and objective way, Kinematics adopts the common subject-object philosophical approach, where the subject is merely an \textsb{observer}\index{observer} and the object the thing observed. In the act of kinematically describing a mechanical event, the observer makes use of two fundamental concepts: position and time. In Newtonian or Classical Mechanics, which is our concern here, \textsb{space}\index{space!physical}, mathematically modeled as a three dimensional Euclidean space, is the context or the framework that enables the subject to perform its positional identifications, when either a certain region of this space or a specific physical entity are the observed objects. The biggest physical space, which includes the whole universe, is defined to be \textsb{absolute}\index{space!absolute}, that is, it has a fixed size, it is completely at rest and not affected by its elements, being a totally independent entity. Any observer attached to the absolute space is also called absolute\index{observer!absolute}. 

When performing a certain measurement on the object under observation, the observer notes that changes may occur: given a point in space or a material element of a physical entity, there may be a set of different measurement values on this point or element that develop one after another, sequentially. Time is the concept that enables the observer to describe these changes by identifying which measurement value follows the other in such a way as to build a set of successive measurement values for every observed point or element.  There is always a specific instance or value of time, called \textsb{instant}\index{instant!of time}, that can label a certain measurement value; in other words, time is infinite and infinitely divisible, or an infinite unlimited complete set, conveniently modeled as a real field. When a limited portion of this set is selected to describe a certain mechanical event, we call this portion a \textsb{period}. If the observer measures the positions of the material elements of a limited physical entity, a biunivocal relation of these elements with space points in a given instant is called
a \textsb{configuration}\index{configuration} of this physical entity and the set of these space points is said to be a \textsb{shape}\index{shape} of this entity. Moreover, a set of these shapes on a certain period of time is defined to be a motion of the physical entity. Still in the context of Newtonian Mechanics, time is considered to be unstoppable, always progressing indefinitely, and the rate of this progress is the same for any observer: time is also considered absolute. Therefore, we can say that every observer's clock is synchronized with a unique reference clock. As a consequence of the ever increasing Newtonian time, an observer describing a certain mechanical event can never label two distinct measurement values relative to the same point of an object by the same instant: simultaneity here makes sense only for different points. Moreover, an observer may eventually become an object when another observer can describe its motion. That said, an observer at rest or moving with constant velocity relative to an absolute observer is called \textsb{inertial}\index{observer!inertial}: it is only from the point of view of an inertial observer that Newton's Law of Inertia is valid.   

Following Hermann Weyl, ``\emph{space and time are commonly regarded as the forms of existence of the real world, \textsb{matter}\index{matter} as its substance.}\footnote{See \aut{Weyl}\cite{weyl_1952_2}, p.1.}'' The purpose of the Newtonian concept of \textsb{mass}\index{mass} is to measure the quantity of this substance, as a tangible entity. Then, we impose that every limited portion of matter, called a body, must have a certain mass. Moreover, it is observed that what constitutes matter, the specific material it is made of, influences its mechanical behavior. In this sense, a mechanical description concerned with material peculiarities is called constitutive. For the purposes of this book, matter is considered to be impenetrable, that is, two simultaneous configurations can never occupy the same portion of space, either totally or partially. A popular topic on the study of matter is its structure, that is, its building blocks, commonly called particles, and the way they are organized. For some reason, this popularity promoted the preconception, even in academic circles, that a physical theory which seeks to describe the overall behavior of matter becomes more reliable, or is much closer to reality, when it considers structural phenomena. In our opinion, this idea tries to attach what it is not attachable. A mechanical theory, expressed as a mathematical model, is good if its results are sufficiently close to experimental data, when available, and if it describes an ample set of physical phenomena, regardless whether structural variables are considered or not. In studying the mechanical behavior of matter, the approach called Continuum Mechanics, which considers a body a continuum, presents both of these virtues, notwithstanding it totally disregards structural phenomena. 

In the context of Continuum Mechanics, where the physical space is a three-dimensional Euclidean affine space, time is a real field and body is a continuum, we conclude that a configuration of a given body is a bijection -- labeled by a real scalar representing an instant of time -- that maps the body to a shape, which is a bounded subset of the three-dimensional Euclidean affine space. In this sense, we can say that the body is a formless portion of matter and only acquires geometric structure, and thus \textsb{volume}, through configuration. Considering the biunivocal property of the configuration and since every portion of matter has mass, we can define on an arbitrary shape of a given body a mass density functional, whose volume integral results the mass of the body relative to the shape in question. For the purposes of our study, we impose that the mass of the body is configuration independent, that is, the mass calculated on any arbitrary shape is always the mass of the body\footnote{In motions where a portion of the mass of a body is burned in order to propel another portion, the above restriction is still valid because any shape of this body as a whole must include not only the propelled portion of its mass, but also that which is burned.}.

O rato roreu a roupa do rei de roma




