\begin{thebibliography}{99.}

\bibitem{backus_1997_2}\aut{Backus, G.} \nob{ Continuum Mechanics.} Disponível em:
<\url{http:// samizdat.mines.edu/backus/}>, Acesso em: fev 2004.

\bibitem{benvenuto_1991}\aut{Benvenuto, E.} \nob{An Introduction to the History of Structural Mechanics, Part I: Statics and Resistance of Solids}. 1. ed. New York: Springer-Verlag, 1991.

\bibitem{brandon_2003_2}\aut{Brandon, R.} \nob{ Large Deformation Kinematics.} Disponível em:
<\url{http://www.me.unm.edu/~rmbrann/gobag.html}>, Acesso em: jan 2007.

\bibitem{chadwick_1999_1}\aut{Chadwick, P.} \nob{ Continuum Mechanics: Concise Theory and Problems.}
1. ed. Mineola: Dover Publications, 1999.

\bibitem{ciarlet_1988_2}\aut{Ciarlet, P. G.} \nob{ Mathematical Elasticity - Volume 1:
Three Dimensional Elasticity.} 1. ed. Amsterdã: Elsevier Science Publishers BV, 1988.

\bibitem{corben_1994_2}\aut{Corben, H. C.; Stehle, P.} \nob{ Classical Mechanics.} 2. ed. New York: Dover Publications, 1994.

\bibitem{dugas_1988_1}\aut{Dugas, R.} \nob{A History of Mechanics.} 1. ed. New York: Dover Publications, 1988.

\bibitem{galileo_1954_2}\aut{Galileo Galilei.} \nob{Dialogs Concerning Two New Sciences.} 1. ed. New York: Dover Publications, 1954.

\bibitem{gurtin_1981_2}\aut{Gurtin, M. E.} \nob{ An Introduction to Continuum Mechanics.} 1. ed. San Diego: Academic Press, 1981.

\bibitem{hanyga_1985_2}\aut{Hanyga, A. } \nob{ Mathematical Theory of Non-Linear Elasticity.} Translated by
R. W. Ogden. 1. ed. Chichester/New York/Warsaw: Elis Horwood Limited/John Wiley \&
Sons/PWN-Polish Scientific Publishers, 1985.

\bibitem{haupt_2002_2}\aut{Haupt, P.} \nob{ Continuum Mechanics and Theory of Materials.} 2. ed. Berlim: Springer-Verlag, 2002.

\bibitem{heilbron_2010_1}\aut{Heilbron, J.L.} \nob{Galileo.} 1. ed. New York: Oxford University Press, 2010.

\bibitem{higdon_1981_3}\aut{Higdon, A. et al.} \nob{Mecânica dos Materiais.} 3. ed. Rio de Janeiro: Guanabara Dois, 1981.

\bibitem{hill_1978_2}\aut{Hill, R.} Aspects of Invariance in Solid Mechanics. \nob{ Advances in Applied Mechanics.} New York: Academic Press, vol. 18, pp. 1-75, 1978.


\bibitem{love_1944_2}\aut{Love, A. E. H.} \nob{ A Treatise on the Mathematical Theory of Elasticity.} 4. ed. New York: Dover Publications, 1944.

\bibitem{lubarda_2002_2}\aut{Lubarda, V. A.} \nob{ Elastoplasticity Theory.} 1. ed. Boca Raton: CRC Press,   2002.

\bibitem{lund_2000_1}\aut{Lund, J.; Byrne, J.} \textit{Leonardo Da Vinci's Tensile Strength Tests: Implications For The Discovery Of Engineering Mechanics} in \nob{Civil. Eng. and Env. Syst.} Vol. 00, 2000, pp. 1-8.


\bibitem{malvern_1969_2}\aut{Malvern, L. E.} \nob{ Introduction to the Mechanics of a Continuous Medium.}
1. ed. Englewood Cliffs: Prentice-Hall, 1969.

\bibitem{marsden_1994_2}\aut{Marsden, J. E.; Hughes, T. J. R.} \nob{ Mathematical
Foundations of Elasticity.} 1. ed. Mineola: Dover Publications, 1994.


\bibitem{maugin_1993_2}\aut{Maugin, G.} \nob{ Material Inhomogeneities in Elasticity.} 1. ed. New York: Chapman \& Hall, 1993.


\bibitem{maxwell_1991_2}\aut{Maxwell, J. C.} \nob{ Matter and Motion.} 1. ed. New York: Dover Publications, 1991.

\bibitem{noll_1974_2}\aut{Noll, W.} \nob{ The Foundations of Mechanics and Thermodynamics: Selected Papers.} 1. ed. Berlim: Springer-Verlag, 1974.

\bibitem{ogden_1997_2}\aut{Ogden, R. W.} \nob{ Non-Linear Elastic Deformations.} 1. ed. Mineola: Dover Publications, 1997.

\bibitem{podio_2000_2}\aut{Podio-Guidugli, P.} A Primer on Elasticity. \nob{ Journal of Elasticity.} Berlim: Springer-Verlag, vol. 40, pp. 1-104, 2000.

\bibitem{popov_1990_1}\aut{Popov, E.} \nob{ Engineering Mechanics of Solids.} 1. ed. Englewood Cliffs: Prentice Hall, 1990.

\bibitem{prager_1997_2}\aut{Prager, W.} \nob{ Introduction to Mechanics of Continua.} 1. ed. Mineola: Dover Publications, 2004.

\bibitem{sedov_1997_2}\aut{Sedov, L. I.} \nob{ Mechanics of Continuous Media - Volume 1.} Translated by J. P. Nowacki. 1. ed. Farrer Road: World Scientific Publishing Co., 1997.

\bibitem{silhavy_1997_2}\aut{\v{S}ilhavý, M.} \nob{ The Mechanics and Thermomechanics of Continuous Media.} 1. ed. Berlim: Springer-Verlag, 1997.

\bibitem{sokolnikoff_1983_2}\aut{Sokolnikoff, I. S.} \nob{ Mathematical Theory of Elasticity.}
2. ed. Malabar: Robert Krieger Publishing Company, 1983.

\bibitem{talpaert_2002_2}\aut{Talpaert, Y. R.} \nob{ Tensor Analysis and Continuum Mechanics.} 1. ed. Dordrecht: Kluwer Academic Publishers, 2002.

\bibitem{teeman_2005_2}\aut{Temam, R. M.; Miranville, A. M.} \nob{ Mathematical Modeling in Continuum Mechanics.} 2. ed. Cambridge: Cambridge University Press, 2005.

\bibitem{timoshenko_1983_1}\aut{Timoshenko, S.} \nob{History of Strength of Materials.} 1 ed. Mineola: Dover Publications, 1983.

\bibitem{truesdell_1968}\aut{Truesdell, C. A.} \nob{Essays in the History of Mechanics}. 1. ed. New York: Springer-Verlag, 1968.

\bibitem{truesdell_1977_2}\aut{Truesdell, C. A.} \nob{ A First Course in Rational Continuum Mechanics - Volume 1}. 1. ed. New York: Academic Press, 1977.


\bibitem{truesdellNoll_1997_2}\aut{Truesdell, C. A.; Noll, W.} \nob{ The Non-Linear Field Theories
of Mechanics.} 3. ed. New York: Springer-Verlag, 2003.

\bibitem{wallace_1998_1}\aut{Wallace, W.} \textit{Galileo's Pisan studies in science and philosophy} in \nob{ The Cambridge Companion to Galileo.} 1. ed. Cambridge: Cambridge University Press, 1998, pp. 27-52.


\bibitem{wangTruesdell_1973_2}\aut{Wang, C.-C; Truesdell, C. A.} \nob{ Introduction to Rational Elasticity.} 1. ed. Leyden: Noordhoff International Publishing, 1973.


\bibitem{weyl_1952_2}\aut{Weyl, H.} \nob{ Space Time Matter.} 4. ed. Mineola: Dover Publications, 1952.

\bibitem{zammattio_1980}\aut{Zammattio, C. et al.} \nob{Leonardo The Scientist.} 1. ed. New York: McGraw-Hill, 1980.



\end{thebibliography}
