\chapter{Introduction to Continuum Mechanics}\label{ch:Collect}


This chapter starts with some historical aspects of Continuum Mechanics in order to situate the reader on the subject, which is a fundamental part of the science of Mechanics as a whole. By history of Continuum Mechanics we mean the direct contributions to the study of motion, strength and deformation of non rigid bodies; in other words, contributions to the mechanics of deformable bodies. Here, it is sufficient to rely on intuitive notions of all these concepts, which will be rigorously defined in the following chapters. Since our description is focused on non rigid bodies, works that contributed solely to the mechanics of material points will not be covered; otherwise, our history would be very, very long. Bla, bla, bla... 


\section{History}


An adequate historical account of the main contributors to the subject of Continuum Mechanics must start with the painter Leonardo Da Vinci (1452-1519). Born in the city of Vinci, a comune of Florence, in the italian region of Tuscany, Da Vinci became an apprentice, at the age of fourteen, in Andrea del Verrocchio's workshop, the most famous florentine artist of the time. Already informally educated on Latin, Geometry and Mathematics, at the workshop, he was taught, besides artistic abilities, a wide range of technical skills, including engineering, architecture, metallurgy and chemistry. Among his notable works, he left not only famous paintings like ``Mona Lisa'' and ``The Last Supper'', but also notebooks, whose descriptive annotations and sketches cover a great variety of themes, from prosaic issues of his everyday life to exercises on Mathematics, architectural drawings, studies on painting and sculpture, engineering, cartography, astronomy, optics, botany, human anatomy, motions of fluids and solids, elastic and inelastic collisions, machine design and so on. Concerning Mechanics, which is our interest here, some scholars on the field contest the boasted scientific value of Da Vinci's works: the most aggressive criticism is given by \aut{Truesdell}\cite{truesdell_1968}, which doubts, in his typical verbosity, whether Da Vinci's sketches and annotations on engineering are really his creation or merely reproduce common technical knowledge of his time. Moreover, Truesdell says that Da Vinci's physical laws, all of them linear, conceived intuitively from simple rules of three, are mostly wrong and the right ones just happened to be correct, since the laws in Physics are either linear or nonlinear. Anyway, since it is not our purpose to explore the credibility of Da Vinci's, we must inform the reader that the italian painter, in his notebook (CA, 82v-b), presents a study entitled ``Testing The Strength of Iron wires of Various Lengths''

 De qualquer forma, despite the ferocious criticism of \aut{Truesdell}\cite{truesdell_1968}, which claims, in his typical verbosity, that Da Vinci's sketches and annotations on Mechanics merely reproduce common technical knowledge of his time, it is important to inform the independent reader that the italian painter, in his notebook (CA, 82v-b), presents a study entitled ``Testing The Strength of Iron wires of Various Lengths'' and falar de barras.     